---
title: "Delegation_Analysis"
author: "Brandon Williams"
date: "2025-07-08"
output: pdf_document
---

# Access packages and data

```{r setup, include=FALSE}
# Pick file path
wd <- "C:/Users/BJW95/OneDrive - University of Pittsburgh/Documents/Projects/Delegation"
# wd <- "C:/Users/bjwil/OneDrive - University of Pittsburgh/Documents/Projects/Delegation"
setwd(wd)

# List of CRAN packages
cran_packages <- c(
  "foreign", 
  "plm", 
  "stargazer", 
  "patchwork", 
  "ks", 
  "lmtest",
  "dplyr", 
  "stringr", 
  "tidyr", 
  "ggplot2", 
  "svglite", 
  "AER",
  "estimatr",
  "ggrepel",
  "modelsummary"
)

# Load all packages
lapply(c("gmoTree", cran_packages), library, character.only = TRUE)

# Load in data
data <- readRDS("Data/deleg_data.rds")
lottery <- readRDS("Data/deleg_lottery.rds")
robot <- readRDS("Data/deleg_robot.rds")
dictator <- readRDS("Data/deleg_dictator.rds")

# Theme for all plots

nice_blue <- "#darkseagreen3"
nice_theme <- theme_classic(base_size = 12) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 13),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),
    axis.line = element_line(color = "black"),
    panel.grid.major.y = element_line(color = "gray90", size = 0.3),
    panel.grid.minor = element_blank()
  )

# Function to generate plots

make_offer_plot <- function(df, chat, delegation, urn_label) {
  df_plot <- df %>%
    filter(urn == urn_label, t > 10, role == "Seller",
           chat == !!chat, delegation == !!delegation) %>%
    count(minX) %>%
    mutate(prop = n / sum(n))
  
  label_main <- if (delegation) "Delegation" else "Take-it-or-Leave-It"
  y_label <- if (delegation) "Minimum Offer" else "Offer"
  title <- paste(label_main, ":", urn_label)
  
  ggplot(df_plot, aes(x = minX, y = prop)) +
    geom_col(fill = "#008080", color = "black", width = 1) +
    scale_x_continuous(breaks = 0:9, limits = c(0, 9)) +
    scale_y_continuous(limits = c(0, 0.65)) +
    xlab(y_label) + ylab("Proportion") +
    ggtitle(title) + nice_theme
}
```

Subject counts (N)

```{r}

subject_counts <- data %>%
  distinct(i, delegation, chat) %>%  
  group_by(delegation, chat) %>%
  summarise(n_subjects = n(), .groups = 'drop')

as.data.frame(subject_counts)
```

Average choices

```{r}
avg_data <- data %>% 
  group_by(delegation, chat, urn) %>% 
  summarise(
    avg_choice = mean(minX)
  )
avg_data
```

Plots (No Chat)

```{r}
# Generate no-chat plots
p1 <- make_offer_plot(data, chat = FALSE, delegation = FALSE, urn_label = "Low")
p2 <- make_offer_plot(data, chat = FALSE, delegation = FALSE, urn_label = "Middle")
p3 <- make_offer_plot(data, chat = FALSE, delegation = FALSE, urn_label = "High")
p4 <- make_offer_plot(data, chat = FALSE, delegation = TRUE, urn_label = "Low")
p5 <- make_offer_plot(data, chat = FALSE, delegation = TRUE, urn_label = "Middle")
p6 <- make_offer_plot(data, chat = FALSE, delegation = TRUE, urn_label = "High")

DrawHist <- (p1 | p2 | p3) / (p4 | p5 | p6)
DrawHist
```
Plots (Chat)

```{r}
# Generate chat-enabled plots
p1c <- make_offer_plot(data, chat = TRUE, delegation = FALSE, urn_label = "Low")
p2c <- make_offer_plot(data, chat = TRUE, delegation = FALSE, urn_label = "Middle")
p3c <- make_offer_plot(data, chat = TRUE, delegation = FALSE, urn_label = "High")
p4c <- make_offer_plot(data, chat = TRUE, delegation = TRUE, urn_label = "Low")
p5c <- make_offer_plot(data, chat = TRUE, delegation = TRUE, urn_label = "Middle")
p6c <- make_offer_plot(data, chat = TRUE, delegation = TRUE, urn_label = "High")

DrawHistc <- (p1c | p2c | p3c) / (p4c | p5c | p6c)
DrawHistc
```

Cumulative offers in Delegation 
```{r, warning=FALSE}
# Helper function to compute proportions of values available in [minX:maxX]
get_value_availability <- function(df, urn_label, chat_cond) {
  df_filtered <- df %>%
    filter(
      delegation == TRUE,
      role == "Seller",
      t > 10,
      urn == urn_label,
      chat == chat_cond
    ) %>%
    select(i, t, minX, maxX, y)
  
  total_offers <- nrow(df_filtered)
  
  expanded <- df_filtered %>%
    rowwise() %>%
    mutate(offer_values = list(minX:maxX)) %>%
    unnest(offer_values) %>%
    ungroup() %>%
    distinct(i, t, offer_values) %>%
    count(offer_values) %>%
    mutate(proportion = n / total_offers)
  
  # Fill missing values from 0 to 9 with 0
  full <- tibble(offer_values = 0:9) %>%
    left_join(expanded, by = "offer_values") %>%
    mutate(proportion = replace_na(proportion, 0))
  
  # Return a list with both the data and label
  return(list(data = full))
}


# Generate 6 datasets
plot_data <- list(
  low_nochat  = get_value_availability(data, "Low", FALSE),
  mid_nochat  = get_value_availability(data, "Middle", FALSE),
  high_nochat = get_value_availability(data, "High", FALSE),
  low_chat    = get_value_availability(data, "Low", TRUE),
  mid_chat    = get_value_availability(data, "Middle", TRUE),
  high_chat   = get_value_availability(data, "High", TRUE)
)

# Plotting function
make_availability_plot <- function(data_list, title) {
  df <- data_list$data
  prop_y0_label <- data_list$label
  
  ggplot(df, aes(x = offer_values, y = proportion)) +
    geom_col(fill = "#008080", color = "black", width = 0.8) +
    scale_x_continuous(breaks = 0:9, limits = c(0, 9)) +
    scale_y_continuous(limits = c(0, 1)) +
    xlab("Value Offered") + ylab("Proportion of Offers Including Value") +
    ggtitle(title) +
    annotate("text", x = 7.5, y = 0.9, label = prop_y0_label, size = 4.5, hjust = 1) +
    nice_theme
}


# Generate plots
p1 <- make_availability_plot(plot_data$low_nochat, "Low Urn (No Chat)")
p2 <- make_availability_plot(plot_data$mid_nochat, "Middle Urn (No Chat)")
p3 <- make_availability_plot(plot_data$high_nochat, "High Urn (No Chat)")
p4 <- make_availability_plot(plot_data$low_chat, "Low Urn (Chat)")
p5 <- make_availability_plot(plot_data$mid_chat, "Middle Urn (Chat)")
p6 <- make_availability_plot(plot_data$high_chat, "High Urn (Chat)")

# Combine with patchwork
DelegationValueAvailability <- (p1 | p2 | p3) / (p4 | p5 | p6)

# Show all
DelegationValueAvailability
```

Inefficiency Types 

```{r}
#### Inefficiency Decomposition ####

ineff_decomp <- data %>%
  filter(role == "Buyer", t > 10) %>%
  group_by(delegation, chat) %>%
  summarise(
    total_n = n(),
    breakdowns = sum(y == 0),
    suboptimal_choices = sum(y > 0 & y < theta, na.rm = TRUE),
    efficient_choices = sum(y >= theta, na.rm = TRUE),
    
    breakdown_rate = breakdowns / total_n,
    suboptimal_rate = suboptimal_choices / total_n,
    efficiency_rate = efficient_choices / total_n,
    
    .groups = "drop"
  ) %>%
  mutate(
    Mech = ifelse(delegation, "Delegation", "TIOLI"),
    Comm = ifelse(chat, "Chat", "NoChat")
  ) %>%
  select(Mech, Comm, breakdown_rate, suboptimal_rate, efficiency_rate)

# Reshape for plotting
ineff_long <- ineff_decomp %>%
  pivot_longer(cols = c(breakdown_rate, suboptimal_rate, efficiency_rate),
               names_to = "Outcome", values_to = "Rate")

ggplot(ineff_long, aes(x = interaction(Mech, Comm), y = Rate, fill = Outcome)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(
    values = c(
      "efficiency_rate" = "#D5D8DC",      # Teal Green
      "suboptimal_rate" = "#008080",      # Muted Orange
      "breakdown_rate"  = "#E76F51"       # Dusty Purple
    ),
    labels = c(
      "efficiency_rate" = "Efficient (y ≥ θ)",
      "suboptimal_rate" = "Suboptimal (0 < y < θ)",
      "breakdown_rate"  = "Breakdown (y = 0)"
    )
  ) +
  labs(
    title = "Decomposition of Buyer Decisions by Treatment",
    x = "Treatment",
    y = "Proportion of Rounds",
    fill = "Decision Outcome"
  ) +
  theme_classic(base_size = 13) +
  scale_x_discrete(labels = function(x) gsub("\\.", "\n", x))
```
Breakdowns 

```{r}
breakdown_data <- data %>%
  filter(role == "Buyer", t > 10) %>%
  group_by(delegation, chat, urn) %>%
  summarise(
    n = n(),
    breakdowns = sum(y == 0),
    breakdowns_rate = mean(y == 0),
    breakdowns_se = sqrt(breakdowns_rate * (1 - breakdowns_rate) / n),
    inefficiencies = sum(y < theta, na.rm = TRUE),
    inefficiency_rate = mean(y < theta, na.rm = TRUE),
    .groups = "drop"
  )
breakdown_data$urn <- factor(breakdown_data$urn, levels = c("High", "Middle", "Low"))

breakdown_data$Mech <- ifelse(breakdown_data$delegation, "Delegation", "TIOLI")
breakdown_data$Comm <- ifelse(breakdown_data$chat, "Chat", "NoChat")

breakdown_table <- pivot_wider(breakdown_data[, c("Mech","Comm",'urn','breakdowns_rate')], names_from = urn, values_from = breakdowns_rate)

breakdown_table[, c("Mech","Comm",'Low','Middle','High')]
```

Efficiency by urn x mechanism

```{r}

ineff_table<- pivot_wider(breakdown_data[, c("Mech","Comm",'urn','inefficiency_rate')], names_from = urn, values_from = inefficiency_rate)

ineff_table[, c("Mech","Comm",'Low','Middle','High')]

# Reshape the inefficiency data to long format
ineff_long <- ineff_table[, c("Mech", "Comm", "Low", "Middle", "High")] %>%
  pivot_longer(cols = c("Low", "Middle", "High"),
               names_to = "Urn", values_to = "Inefficiency") %>%
  mutate(
    Urn = factor(Urn, levels = c("High", "Middle", "Low")),
    Treatment = interaction(Mech, Comm)
  )

# Plot
ggplot(ineff_long, aes(x = Treatment, y = Inefficiency, fill = Urn)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7, color = "black") +
  scale_fill_manual(
    values = c(
      "Low" = "#D5D8DC",     # Light gray
      "Middle" = "#008080",  # Teal
      "High" = "#E76F51"     # Burnt coral / high contrast
    )
  ) +
  labs(
    title = "Inefficiency Rate by Treatment and Urn",
    x = "Treatment (Delegation × Chat)",
    y = "Inefficiency Rate",
    fill = "Urn Type"
  ) +
  scale_x_discrete(labels = function(x) {
    sapply(strsplit(as.character(x), "\\."), function(p) paste0(p[1], "\n", p[2]))
  }) +
  theme_classic(base_size = 13)
```

Who gains the most through each mechanism / urn? 

```{r}

# Payoffs

payoff_summary <- data %>%
  filter(t > 10) %>%
  mutate(
    Mech = ifelse(delegation, "Delegation", "TIOLI"),
    Comm = ifelse(chat, "Chat", "NoChat")
  ) %>%
  group_by(Mech, Comm, urn, role) %>%
  summarise(avg_pay = mean(pay, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = role,
    values_from = avg_pay,
    names_prefix = "avg_"
  ) %>%
  mutate(
    Shape = paste(Mech, "–", Comm),
    urn = factor(urn, levels = c("Low", "Middle", "High"))
  )

payoff_summary <- payoff_summary %>%
  mutate(
    Shape = paste(Mech, "–", Comm),
    urn = factor(urn, levels = c("Low", "Middle", "High")),
    label = paste(Mech, "–", Comm, "–", urn)  # Full label
  )

# Shapes for Mech × Comm
mech_comm_shapes <- c(
  "TIOLI – NoChat"      = 16,  # filled circle
  "TIOLI – Chat"        = 17,  # filled triangle
  "Delegation – NoChat" = 15,  # filled square
  "Delegation – Chat"   = 18   # filled diamond
)

# Colors for Urn types
urn_colors <- c(
  "Low"    = "#7D7D7D",   # Slate gray
  "Middle" = "#008080",   # Teal
  "High"   = "#E76F51"    # Burnt coral
)

ggplot(payoff_summary, aes(x = avg_Seller, y = avg_Buyer)) +
  geom_point(aes(color = urn, shape = Shape), size = 4) +
  geom_text_repel(aes(label = label), size = 3.5, max.overlaps = 100) +
  scale_color_manual(values = urn_colors, name = "Urn Type") +
  scale_shape_manual(values = mech_comm_shapes, name = "Mechanism × Communication") +
  labs(
    title = "Average Buyer vs Seller Payoffs by Treatment and Urn",
    x = "Seller Payoff",
    y = "Buyer Payoff"
  ) +
  theme_classic(base_size = 13) +
  theme(
    legend.position = "right",
    legend.box = "vertical"
  )

```

Lottery data

```{r}
# Average lottery selection by 

lottery_avg <- lottery %>%
  group_by(delegation, chat, urn) %>%
  summarise(
    avgLotteryX = mean(lotteryMinX[ifelse(type == "Delegation", delegation, !delegation)], na.rm = TRUE),
    neutralMinX = mean(neutralMinX),
    .groups = "drop"
  )
lottery_avg

```
Robot optimization data

```{r}

robot_avg <- robot %>%
  group_by(urn, chat, delegation) %>%
  summarise(
    avgRobotX = mean(minX, na.rm = TRUE),
  )
robot_avg

```

Other regarding choices

```{r}

dictator_avg <- dictator %>%
  group_by(delegation, chat, urn) %>%
  summarise(
    avgDictatorX = mean(action, na.rm = TRUE),
    selfish = mean(maxAction)
  )

dictator_avg

```

Comparison of averages 

```{r}
merged_data <- avg_data %>%
  left_join(robot_avg %>% select(delegation, chat, urn, avgRobotX), 
            by = c("delegation", "chat", "urn")) %>%
  left_join(lottery_avg %>% select(delegation, chat, urn, avgLotteryX), 
            by = c("delegation", "chat", "urn")) %>%
  left_join(dictator_avg %>% select(delegation, chat, urn, avgDictatorX), 
            by = c("delegation", "chat", "urn")) %>%
  mutate(
    urn = factor(urn, levels = c("Low", "Middle", "High"))
  ) %>%
  arrange(delegation, chat, urn)
merged_data

```

```{r}
# 1. Filter chat == FALSE
filtered_data <- merged_data %>%
  filter(chat == FALSE)

# 2. Reshape to long format
long_data <- filtered_data %>%
  pivot_longer(
    cols = c(avg_choice, avgRobotX, avgLotteryX, avgDictatorX),
    names_to = "source",
    values_to = "value"
  ) %>%
  mutate(
    source = factor(source, levels = c("avg_choice", "avgRobotX", "avgLotteryX", "avgDictatorX"))
  )

# 3. Plot with facet for delegation
ggplot(long_data, aes(x = source, y = value, group = urn, color = urn)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  facet_wrap(~ delegation, labeller = label_both) +
  labs(
    title = "Average (Minimum) Offers (No chat only)",
    x = "Source",
    y = "(Minimum) Offer"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Regression analysis of offers

```{r results='asis', message=FALSE, warning=FALSE}



seller <- subset(data, role == "Seller")
seller$dist <- as.factor(seller$urn)

lm_minX <- lm_robust(minX ~ dist + delegation * chat, data = seller, cluster = i)

modelsummary(lm_minX, stars = TRUE)

# LATEX output 
# modelsummary(lm_minX, stars = TRUE, output = "latex")

# Obsididan ouput

# Create human-readable names
pretty_names <- c(
  "(Intercept)" = "Intercept",
  "distLow" = "Low Urn",
  "distMiddle" = "Middle Urn",
  "delegationTRUE" = "Delegation",
  "chatTRUE" = "Chat",
  "delegationTRUE:chatTRUE" = "Delegation × Chat"
)

# Tidy + stars + label mapping
tidy_lm <- tidy(lm_minX) %>%
  mutate(
    stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.1   ~ "+",
      TRUE            ~ ""
    ),
    estimate_star = paste0(round(estimate, 3), stars),
    std.error = round(std.error, 3),
    p.value = signif(p.value, 3),
    term = pretty_names[term]  # rename terms
  ) %>%
  select(term, estimate_star, std.error, p.value)

# Print table to Obsidian-compatible markdown
kable(
  tidy_lm,
  format = "markdown",
  col.names = c("Term", "Estimate", "Std. Error", "p-value")
)




```


Include full interaction

```{r results='asis', message=FALSE, warning=FALSE}

seller <- subset(data, role == "Seller")
seller$dist <- as.factor(seller$urn)

lm_minX_full <- lm_robust(minX ~ dist * delegation * chat, data = seller, cluster = i)

modelsummary(lm_minX_full, stars = TRUE)

# LATEX output 
# modelsummary(lm_minX, stars = TRUE, output = "latex")

# Obsididan ouput

# Create human-readable names
pretty_names <- c(
  "(Intercept)" = "Intercept",
  "distLow" = "Low Urn",
  "distMiddle" = "Middle Urn",
  "delegationTRUE" = "Delegation",
  "chatTRUE" = "Chat",
  "distLow:delegationTRUE" = "Low Urn x Delegation",
  "distMiddle:delegationTRUE" = "Middle Urn x Delegation",
  "distLow:chatTRUE" = "Low Urn x Chat",
  "distMiddle:chatTRUE" = "Middle Urn x Chat",
  "delegationTRUE:chatTRUE" = "Delegation × Chat",
  "distLow:delegationTRUE:chatTRUE" = "Low Urn x Delegation × Chat",
  "distMiddle:delegationTRUE:chatTRUE" = "Middle Urn x Delegation × Chat"
)

# Tidy + stars + label mapping
tidy_lm <- tidy(lm_minX_full) %>%
  mutate(
    stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.1   ~ "+",
      TRUE            ~ ""
    ),
    estimate_star = paste0(round(estimate, 3), stars),
    std.error = round(std.error, 3),
    p.value = signif(p.value, 3),
    term = pretty_names[term]  # rename terms
  ) %>%
  select(term, estimate_star, std.error, p.value)

# Print table to Obsidian-compatible markdown
kable(
  tidy_lm,
  format = "markdown",
  col.names = c("Term", "Estimate", "Std. Error", "p-value")
)




```


Fitted values

```{r}

new_data <- expand.grid(
  dist = factor(c("Low", "Middle", "High"), levels = levels(seller$dist)),
  delegation = c(TRUE, FALSE),
  chat = c(TRUE, FALSE)
)

preds <- predict(lm_minX_full, newdata = new_data, se.fit = TRUE)

new_data <- new_data %>%
  mutate(
    fit = preds$fit,
    se = preds$se.fit,
    lower = fit - 1.96 * se,
    upper = fit + 1.96 * se
  )

new_data

new_data$dist <- factor(new_data$dist, levels = c("Low", "Middle", "High"))

ggplot(new_data, aes(x = dist, y = fit, color = as.factor(delegation), group = delegation)) +
  geom_point(size = 3) +
  geom_line() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  facet_wrap(~ chat, labeller = label_both) +
  labs(
    x = "Urn Type",
    y = "Predicted minX",
    color = "Delegation",
    title = "Predicted minX by Urn and Delegation (Faceted by Chat)"
  ) +
  theme_minimal()
```

Regression on outcome choices

```{r results='asis', message=FALSE, warning=FALSE}

lm_y <- lm_robust(y ~ dist + delegation * chat, data = seller, cluster = i)

modelsummary(lm_y, stars = TRUE)

# LATEX output 
# modelsummary(lm_y, stars = TRUE, output = "latex")

# Obsididan ouput

# Create human-readable names
pretty_names <- c(
  "(Intercept)" = "Intercept",
  "distLow" = "Low Urn",
  "distMiddle" = "Middle Urn",
  "delegationTRUE" = "Delegation",
  "chatTRUE" = "Chat",
  "delegationTRUE:chatTRUE" = "Delegation × Chat"
)

# Tidy + stars + label mapping
tidy_lm <- tidy(lm_y) %>%
  mutate(
    stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.1   ~ "+",
      TRUE            ~ ""
    ),
    estimate_star = paste0(round(estimate, 3), stars),
    std.error = round(std.error, 3),
    p.value = signif(p.value, 3),
    term = pretty_names[term]  # rename terms
  ) %>%
  select(term, estimate_star, std.error, p.value)

# Print table to Obsidian-compatible markdown
kable(
  tidy_lm,
  format = "markdown",
  col.names = c("Term", "Estimate", "Std. Error", "p-value")
)




```
See who optimizes based on their lottery answer

```{r}

# Create two different data sets, one for each delegation treatment

robot_del <- robot %>% 
  filter(delegation == TRUE) %>% 
  select(i, urn, minX) %>%
  rename(robotX = minX)
robot_tioli <- robot %>% 
  filter(delegation == FALSE) %>% 
  select(i, urn, minX) %>%
  rename(robotX = minX)

# Split the pool of lottery choices by delegation type
# (this will have people from both delegation treatments)

lottery_del <- lottery %>% 
  filter(type == "Delegation") %>%
  select(i, lotteryMinX, urn)
  
lottery_tioli <- lottery %>% 
  filter(type == "TIOLI") %>%
  select(i, lotteryMinX, urn)


# Join (confirmed this matches the subject count times three observations each)

join_del <- inner_join(lottery_del, robot_del, by = c("i", "urn"))
join_tioli <- inner_join(lottery_tioli, robot_tioli, by = c("i", "urn"))

# How many match their risk preferences?

join_del <- join_del %>%
  mutate(
    diff = robotX - lotteryMinX,
    match = robotX == lotteryMinX)

join_tioli <- join_tioli %>%
  mutate(
    diff = robotX - lotteryMinX,
    match = robotX == lotteryMinX)

robot_vs_lottery <- bind_rows(
  join_del %>% mutate(type = "Delegation"),
  join_tioli %>% mutate(type = "TIOLI")
)


ggplot(robot_vs_lottery, aes(x = diff, fill = type)) +
  geom_histogram(
    aes(y = after_stat(count / sum(count))),
    binwidth = 1,
    position = "dodge",
    color = "black"
  ) +
  labs(
    title = "Proportion of Differences Between robotX and lotteryMinX",
    x = "robotX - lotteryMinX",
    y = "Proportion",
    fill = "Condition"
  ) +
  theme_minimal()

match_summary <- robot_vs_lottery %>%
  group_by(i) %>%
  summarise(
    match_count = sum(match, na.rm = TRUE),
    total_urns = n_distinct(urn)
  ) %>%
  filter(total_urns == 3)  # Only include those with all 3 urns

# 3. Create summary table of 0/1/2/3 matches
match_table <- match_summary %>%
  count(match_count) %>%
  arrange(match_count)

# View the table
match_table
```




